SiO Molecular Jets around young stars - A numerical perspective

### 2013-08-12 PAPER/ms_final_bv.tex
Modified paper tex and added new figures

\begin{abstract}
  % context heading (optional)
  {}  
  % aims heading (mandatory)
   {}
  % methods heading (mandatory)
   {}
  % results heading (mandatory)
   {}
  % conclusions heading (optional), leave it empty if necessary 
\end{abstract}

@@

Even with myriad of empirical evidences, the exact nature of SiO and
CO outflow is not clear. For a complete understanding it is imperative
to compliment these observations with theoretical models. Many models
based on hydro-simulations and steady state shock calculations
were proposed to explain the observational signatures of molecular
outflows. Among them the two main models are that of wide-angled wind
driven \cite{Shu:1991p14071} and jet driven outflow
\cite{Canto:1991p14123}. 
The most popular among them is the jet driven model as wind
driven molecular outflows not only fail to match observed PV digarams \cite{Cabrit:1992p14098}
but also tend to sweep large quantitity of material at the extermities
of the lobes \cite{Masson:1992p14101}. While the jet driven models could sucessfully
derive the global outflow shapes and mass velocity relations of CO
outflows \citep{Raga:1993p9948, Masson:1993p9661} . There have also been some attempts to combine these two
models into one \cite{Shang:2006p14268} to explain the global observational features. 
However, most of these dynamical models do not
account for shock chemsitry. Instead, shock chemistry is studied
independently without dynamics using plane parallel shock
models \citep{Schilke:1997p14140,Flower:2003p11236} . The magneto-hydrodynamic (MHD) calculations by \cite{Glassgold:1991p13703} suggested that molecules
like SiO and CO could as form within the jet. Similar conclusions of
molecules surving in steady state disk winds have also been shown
\cite{Panoglou:2012p10039}. Very limited simulations have modelled the outflow
dynamics including molecular chemistry but in absence of
magnetic fields \citep{Raga:1995p12965, Smith:2003p9985}.

@@

The jet enters into the medium through a nozzle of radius $R_{\rm jet}$
from the lower boundary (z = 0). The jet density is fixed to
be $10^{5} \rm cm^{-3}$ and it has a radius of $2.5\times10^{15} \rm cm \sim
167\,AU$. The jet is injected into the domain with a typical 
velocity of $v_{\rm jet,0}$ = 100 km/s as is the case for most low mass stellar jets
specially the low velocity component. The constant jet velocity is
superimposed with periodic pulsation of the form,
\begin{equation}
v_{\rm jet} = v_{\rm jet,0}(1.0 + A\,sin(2.0\pi t/T_{\rm p}))
\end{equation}
where the amplitude A is 0.25 and time period $T_{\rm p} = 70$
years. The pressure at the surface of the jet is $10^{-10} \rm
dyne\,\,cm^{-2}$ corresponding to a temperature of $T_{\rm jet} \sim
4\times10^{3}\,K$. Inside the jet beam, a radial variation of thermal
and magnetic pressure is adopted to maintain a magneto-static
equilibrium. We adopt the
same radial profiles used by \cite{Stone:2000p2650} for all our runs. 
Based on these profiles, the toroidal magnetic field is assumed to be
zero at the axis and achieves a maximum at some radius, $r_{\rm m}$
inside the jet. The maximum value, $B_{\phi,\rm m}$, depends on the
plasma $\beta$ which is a parameter used for our study along with the
density contrast $\eta$.
%

@@

\section{Radiative Transfer}
Radiative transfer modeling used for post processing.

\subsection{The radiative transfer code} \label{subsec:radiative_transfer_code}
The radiative transfer program used is LIME (LIne Modeling Engine; \citealt{Brinch:2010p13078}), which  calculates line intensities based on a weighted sample of randomly chosen points in a continuous 3D model. The method of selecting these points is given in section \ref{subsec:gridding}. At each of these points, the density of the main collision partner (equivalent amount of H$_2$, given by n(H$_2$)+ 0.5 n(H)), gas and dust temperatures, velocity, molecular abundances and unresolved turbulent velocity are specified. These points are then smoothed by Lloyd's algorithm \citep{Lloyd1982} in order to minimise the variation in distance between points whilst keeping the same underlying distribution. These points are then connected by Delaunay triangulation and it is between the points connected by this method that photon are allowed to propagate (fig.~\ref{grid}). The level populations of the selected molecules are calculated at each of these points from collisional and radiative (de)excitation and the local radiation field is calculated. This is repeated 20 times with the populations of each level converging towards a single value. This number of iterations is sufficient for the signal to noise ratio of the level populations (as defined in \citealt{Brinch:2010p13078}) to exceed 1000 for 99\% of the points, ensuring that the simulation has converged on a stable level population. After 20 iterations the model is ray-traced in order to produce synthetic brightness maps. The average of ten separate runs was taken to minimise the artefacts in the output images, resulting from the grid construction (Fig.~\ref{points}).

\begin{figure}
 \includegraphics[width=84mm]{\figpath/eta2Beta10.png}
 \caption{A plot of the points selected by the gridding process and the paths down which photons can propagate for points in the r,z plane. The points are color coded by the density distribution (in m$^{-3}$, as used in LIME) and are more concentrated in the high density knots.}
\label{grid}
\end{figure}

@@

\begin{figure}
 \includegraphics[width=63mm]{\figpath/eta2Beta10_histo.png}
 \caption{A 2D histogram of the point distribution throughout the model. The disc and envelope can be seen as two separate entities which have to be sampled using different point distributions.}
 \label{points}
\end{figure}

\subsection{SiO abundance}
Need refs for this!!\smallskip
The amount of SiO is determined by the local velocity and temperature. The fractional abundance is given by the equation:\smallskip

$log(X)=-2.48\times10^{-8}v^5 + 5.50\times10^{-6}v^4 - 4.28\times10^{-4}v^3 + 1.24\times10^{-2}v^2 + 2.52\times10^{-2}v - 1.20\times10^{1}$\smallskip

where v is in kilometres per second. In addition to this if the temperature at the point is greater than 92,000K (the temperature of the Si-O bond disassociation energy) the abundance is reduced to 10$^{-30}$.

@@

\section{Parameter Survey}
\begin{table*}
\centering
\caption{Summary from parameter runs}
\begin{tabular}{l | l | l | l | l | l }
\hline
Run & Cooling Mode & $\eta$ & $\beta$ & Peak Intensity [K] & FWHM at Bow
Shock \\
\hline\hline
adi1010 & Nil (Adiabatic) & 10 & 10 &&\\
pow1010 & Power law & 10 & 10 &&\\
atm1010 & Atomic & 10 & 10 &&\\
atm101 & Atomic & 10 & 1 &&\\
atm210 & Atomic & 2 & 10 &&\\
tab1010 & Tabulated & 10 & 10 &&\\
tab210 & Tabulated & 2 & 10&&\\
mol1010 & Molecular & 10 & 10&&\\
mol510 & Molecular & 3 & 10&&\\
\hline
\end{tabular}
\label{tab:result1}
\end{table*}

\section{Results}
\subsection{Comparison of cooling prescription}
figure nos. 2 -- Similuated images, emission maps for eta = 10.
\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/jetplt1_1010.png}
 \caption{Jet Volume Density for different cooling modes with
   $\eta$ = 10 and $\beta$ = 10.}
\label{grid}
\end{figure*}

\begin{figure}
 \includegraphics[width=84mm]{\figpath/J2-1_mom0maps_cooling.png}
 \caption{A plot of the integrated SiO J2-1 emission from 5 models, 
each using a different method to calculate cooling and all with $\eta$=10 $\beta$=10.}
\label{cooling2-1}
\end{figure}

\subsection{Effects of changing $\eta$ and $\beta$}
figure nos. 2 -- Simulated images describing lengths, collimation and
emission for lower eta more stronger. 
\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/jetplt2_diffetabeta.png}
 \caption{Jet Volume Density for atomic cooling with $\eta$ = 10 but $\beta$ = 1 and
   10 and that with tabulated cooling with $\beta$ = 10 but $\eta$ = 2
   and 10.}
\label{grid}
\end{figure*}

\subsection{Molecular cooling and H$_2$ Chemistry}
In case of molecular outflows, the most appropiate form of cooling
prescribtion among the ones described in sect.~\ref{sec:chem} is
that which involves the evolution of H$_2$ chemistry along with
contributions to cooling from fixed fractions of other molecules like
CO, OH etc. Figure~\ref{fig:Hfracsim} shows the density of
various hydrogen species in the outflow. Also shown in the figure is
the temperature and mean molecular weight $\mu$. As the jet propgates
into the medium

figure nos. 2 --> Explain molecular cooling in more details. With the
mechanism of formation and destruction of HI, H2 and HII using
figures. 
\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/molcjet_310.png}
 \caption{Fraction of hydrogen species in the run with molecular
   cooling having $\eta$ = 3 and $\beta$ = 10.}
\label{fig:Hfracsim}
\end{figure*}

\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/molcjet_310_molform.png}
 \caption{Dependence of hydrogen fractions on the temperature at two
   points in the flow, viz. the interface of the knot with molecular
   medium and at the bow shock}
\label{grid}
\end{figure*}

\subsection{SiO Abundance profile}
figure nos. 2 -- Comparison of three profiles for SiO Abundance,
Contour maps for two transitions for top hat profile.
\begin{figure}
 \includegraphics[width=1.\columnwidth]{\figpath/EM_diffabun.png}
 \caption{Variation of 2-$>$1 SiO emission for runs with molecular
   cooling having $\eta$ = 3 and $\beta$ = 10 and different abundance profiles.}
\label{grid}
\end{figure}

\begin{figure}
 \includegraphics[width=1.\columnwidth]{\figpath/contourmapsSiOTopHat.png}
 \caption{Contours of SiO 2-$>$1 and 8-$>$7 emission obtained using the
   SiO Top Hat profile.}
\label{grid}
\end{figure}

\subsection{Spectra and PV diagrams}
figure nos. 1 -- one transition 3 different angles, Comparison of
spectra and line ratios

\begin{figure}
 \includegraphics[width=1.\columnwidth]{\figpath/PV_diffabun.png}
 \caption{Variation of position-velocity diagram for 2-$>$1 SiO emission for runs with molecular
   cooling having $\eta$ = 3 and $\beta$ = 10 and different abundance profiles.}
\label{grid}
\end{figure}

\section{Discussion}
Effects of varying $\eta$\\
Lower $\eta$ gives better emission.

Importance of cooling prescriptions\\
Bulk of emission comes from cooling instabilities.

Effects of varing abundance profile ... Crucial ingredient in
modeling molecular bullets \\

Spectra and PV Digrams\\
Line - withs $\sim$ 15-25 km\,s$^{-1}$. Wedge shaped PV diagrams
a signature feature of jet driven molecular outflows. Counterparts in observations.

Comparison with Observed results \\
Different line transition trace different components of SiO
emission. The low level transitions trace the region where the jet
interacts with molecular medium. The clumpy nature in its emission is
from the turbulent nature of its interaction. The higher transitions
trace a much more collimated region along with the terminal shock. 
As seen by Chandler \& Richer 2001 and Hirano 2006.
\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/linejets_thnolog.png}
 \caption{SiO emission from different line transitions for the run with molecular
   cooling having $\eta$ = 3 and $\beta$ = 10 using top-hat fractional
   abundance profile. }
\label{grid}
\end{figure*}

Advantage of non-LTE modeling over LVG \cite{Cabrit:2007p13804}, \cite{Lee:2008p13697}. 
Line ratios $\sim$ 1 can be explained well for low mass stars and 
\cite{Leurini:2013p13165} for high mass stars. \\
\begin{figure}
 \includegraphics[width=1.\columnwidth]{\figpath/LineRatio_TH_60_Y205.png}
 \caption{{\em{Top}} Line profiles in SiO J = 2-1, 5-4 and 8-7 at one
   the inner knot for the molecular cooling run with $\eta$ = 3 and
   $\beta$ = 10 using top hat abundance profile. The profiles are
   obatined when the angle of inclination is 60$^{\circ}$ with respect
 to line of sight. {\em{Bottom}} Line temperature ratios T(8-7)/T(5-4)
and T(2-1)/T(5-4), as a function of velocity}
\label{grid}
\end{figure}

EHV component ?? \\

Limitations\\

\section{ALMA view}
ALMA view of the reference run and stress of applying our synthetic techniques
to study the molecular outflows in more details.\\
Figure 6. for ALMA -- \\

>>>>

\begin{abstract}
  % context heading (optional)
  {A bipolar, violent, collimated outflow is one of the first signpost of star
formation. Such an outflow is believed to be launched magnetically from
underlying accretion disk. As it propagates through the molecular
cloud, it injects energy and momentum via shocks 
and fosters chemical evolution by forming, destroying and entraining
molecules along its path.}  
  % aims heading (mandatory)
   {High 
velocity molecular outflows are extensively studied for both low mass
and high mass stars. They are usually observed using standard sub-mm outflow and shocks
tracers like CO and SiO respectively. However, the exact nature of
excitation of these molecules is not yet clear due to lack of models that
simultaneously study the dynamics along with complex molecular chemistry.}
  % methods heading (mandatory)
   { For such a study, we have perfomed MHD simulations
of jet propagation into a molecular cloud using the PLUTO
code. Firstly, we evolve the jet dynamical quantities in conjunction
with different non-equilibrium cooling prescriptions of varying complexities. 
The most complex is that of molecular cooling along with
H$_2$ chemistry. This prescription allows us to track the formation
and destruction of HI, HII and H$_2$ along with the flow dynamics. 
The final state of the jet obtained for each cooling model is then
post-processed using a non-LTE radiative line transfer code, LIME,  to
obtain SiO emission maps, spectra and PV digram.}
  % results heading (mandatory)
   {We find that the strength of SiO
emission depends strongly on the cooling prescription and SiO fractional
abundance profile. We find that the bulk of SiO emission comes from
the interface between
the jet and the ambient molecular medium mainly excited to due to
shocks. Further, we see that some SiO can be produced within the jet
specially close to the bow shock due to instabilities associated with
cooling. We have used these emission maps to give predictions for ALMA
observations. Also, our model can very well reproduce the observed 
spectra, line ratios and PV diagrams of young outflows from Class 0 sources.}
  % conclusions heading (optional), leave it empty if necessary 
\end{abstract}

@@

Even with myriad of empirical evidences, the exact nature of SiO and
CO outflow is not clear. For a complete understanding it is imperative
to compliment these observations with theoretical models. Many models
based on hydro-simulations and steady state shock calculations
were proposed to explain the observational signatures of molecular
outflows. Among them the two main models are that of wide-angled wind
driven \cite{Shu:1991p14071} and jet driven outflow
\cite{Canto:1991p14123}. 
The most popular among them is the jet driven model as wind
driven molecular outflows not only fail to match observed PV digarams \cite{Cabrit:1992p14098}
but also tend to sweep large quantitity of material at the extermities
of the lobes \cite{Masson:1992p14101}. While the jet driven models could sucessfully
derive the global outflow shapes and mass velocity relations of CO
outflows \citep{Raga:1993p9948, Masson:1993p9661} . There have also been some attempts to combine these two
models into one \cite{Shang:2006p14268} to explain the global observational features. 
However, most of these dynamical models do not
account for shock chemsitry. Instead, shock chemistry is studied
independently using steady state non-dissociative C-type and
dissociative J-type shocks models \citep{Neufeld:1989p11689, Schilke:1997p14140,Flower:2003p11236}. 
The magneto-hydrodynamic (MHD) calculations by \cite{Glassgold:1991p13703} suggested that molecules
like SiO and CO could as form within the jet. Similar conclusions of
molecules surving in steady state disk winds have also been shown
\cite{Panoglou:2012p10039}. 
Very limited simulations have modelled the outflow
dynamics including molecular chemistry but in absence of
magnetic fields \citep{Raga:1995p12965, Smith:2003p9985}.

@@

The jet enters into the medium through a nozzle of radius $R_{\rm jet}$
from the lower boundary (z = 0). The jet density is fixed to
be $10^{5} \rm cm^{-3}$ and it has a radius of $2.5\times10^{15} \rm cm \sim
167\,AU$. The jet is injected into the domain with a typical 
velocity of $v_{\rm jet,0}$ = 100 km/s as is the case for most low mass stellar jets
specially the low velocity component. The constant jet velocity is
superimposed with periodic pulsation of the form,
\begin{equation}
v_{\rm jet} = v_{\rm jet,0}(1.0 + A\,sin(2.0\pi t/T_{\rm p}))
\end{equation}
where the amplitude A is 0.25 and time period $T_{\rm p} = 70$
years. The pressure at the surface of the jet is $10^{-10} \rm
dyne\,\,cm^{-2}$ corresponding to a temperature of $T_{\rm jet} \sim
4\times10^{3}\,K$. Inside the jet beam, a radial variation of thermal
and magnetic pressure is adopted to maintain a magneto-static
equilibrium. We adopt the
same radial profiles used by \cite{Stone:2000p2650} for all our runs. 
Based on these profiles, the toroidal magnetic field is assumed to be
zero at the axis and achieves a maximum at some radius, $r_{\rm m}$
inside the jet. The maximum value, $B_{\phi,\rm m}$, depends on the
plasma $\beta$ which is kept fixed for all our runs to be at a value
of 10.0. This corresponds to the maximum field strength, $B_{\phi,\rm m}$ $\sim
16\mu\,G$.

%

@@

\section{Radiative Transfer}
\label{sec:radtrans}
Radiative transfer modeling used for post processing.

\subsection{The radiative transfer code} \label{subsec:radiative_transfer_code}
The radiative transfer program used is LIME (LIne Modeling Engine; \citealt{Brinch:2010p13078}), which  calculates line intensities based on a weighted sample of randomly chosen points in a continuous 3D model. The method of selecting these points is given in section \ref{subsec:gridding}. At each of these points, the density of the main collision partner (equivalent amount of H$_2$, given by n(H$_2$)+ 0.5 n(H)), gas and dust temperatures, velocity, molecular abundances and unresolved turbulent velocity are specified. These points are then smoothed by Lloyd's algorithm \citep{Lloyd1982} in order to minimise the variation in distance between points whilst keeping the same underlying distribution. These points are then connected by Delaunay triangulation and it is between the points connected by this method that photon are allowed to propagate (fig.~\ref{grid}). The level populations of the selected molecules are calculated at each of these points from collisional and radiative (de)excitation and the local radiation field is calculated. This is repeated 20 times with the populations of each level converging towards a single value. This number of iterations is sufficient for the signal to noise ratio of the level populations (as defined in \citealt{Brinch:2010p13078}) to exceed 1000 for 99\% of the points, ensuring that the simulation has converged on a stable level population. After 20 iterations the model is ray-traced in order to produce synthetic brightness maps. The average of ten separate runs was taken to minimise the artefacts in the output images, resulting from the grid construction.% (Fig.~\ref{points}).

\begin{figure}
 \includegraphics[width=84mm]{\figpath/eta2Beta10.png}
 \caption{A plot of the points selected by the gridding process and the paths down which photons can propagate for points in the r,z plane. The points are color coded by the density distribution (in m$^{-3}$, as used in LIME) and are more concentrated in the high density knots.}
\label{grid} % This plots has to be changed. 
\end{figure}

@@

\subsection{SiO abundance}
\label{ssec:sioabun}
Need refs for this!!\smallskip
The amount of SiO is determined by the local velocity and
temperature. The fractional abundance is given by the equation:

\begin{eqnarray}
log(X)=-2.48\times10^{-8}v^5 + 5.50\times10^{-6}v^4 \nonumber \\
- 4.28\times10^{-4}v^3 + 1.24\times10^{-2}v^2 \nonumber \\
+ 2.52\times10^{-2}v - 1.20\times10^{1}
\label{eq:fracSio}
\end{eqnarray}

where v is the shock velocity in kilometres per second. 
In addition to this if the temperature at the point is greater than 92,000K (the temperature of the Si-O bond disassociation energy) the abundance is reduced to 10$^{-12}$.

@@

\section{Parameter Survey}
The results and analysis for the present work are divided in two parts,
viz., the dynamical numerical simulations and the post-processing with
radiative transfer code. For each part we have used certain parameters
and the effect of changing them is studied. Ideally all these parameters should come from observational
results. However, not all quantities needed for our study are well
constraint by the observations, thus allowing us vary them as
free parameters. Such a parameter survey provides better handle on the range of allowed
values on qualitative comparison with observations. 
%

For the first part of the study concerning the dynamical numerical
simulations, we will focus on two main parameters. They are the
prescribtion of cooling and the density contrast between the jet and
the ambient medium denoted by $\eta$. The various cooling prescribtions used for the present study are
described in details in section~\ref{sec:chem}. A value of $\eta > 1$, mean that
that jet is overdense with respect to the medium. In all our runs with
different cooling prescribtions, we have assumed the jet to be
overdense by 10 times that of the ambient medium. Additionally, for
the atomic, tabulated and molecular cooling runs we have used a
value of $\eta$ of the order of unity indicating similar densities in
the jet and the medium. 
The magentic field strength is kept to be fixed using $\beta = 10.0$,
for all our runs. Table~\ref{tab:result1} lists all the runs with
varying $\eta$ and cooling prescriptions, along with the peak
intensity and line widths at the bow shock.
%

The post-processing phase is done in details for the molecular run
with $\eta = 3$. To obtain the SiO emission maps and corresponding
spectra, two additional free parameters are required along with other
inputs described in section~\ref{sec:radtrans} . They are
the fractional abundance profile of SiO and the angle of inclination with
respect to line of sight denoted by $\phi$. Section~\ref{ssec:sioabun}
describes all the profiles used for the present study. The obivious
choice for the inclination angle, $\phi$ is 90$^{\circ}$ indicating
the outflow is in the plane of the sky. 
Additionally, we have used two other angles of inclination apart from the
plane of the sky, i.e, $\phi = 45^{\circ}, 60^{\circ}$ to compare
results with observations. The runs with different fractional
abundance profiles are described in table~\ref{tab:result2}

\begin{table*}
\centering
\caption{Summary from parameter runs.}
\begin{tabular}{l | l | l | l | l | l }
\hline
Run & Cooling Mode & $\eta$ & Peak Intensity [K] & $\Delta$v [km s$^{-1}$] \\
\hline\hline
adi1010 & Nil (Adiabatic) & 10 & &\\
pow1010 & Power law & 10 & &\\
atm1010 & Atomic & 10 & &\\
atm210 & Atomic & 2 & &\\
tab1010 & Tabulated & 10 & &\\
tab210 & Tabulated & 2 & &\\
mol1010 & Molecular & 10 & &\\
mol310 & Molecular & 3 & &\\
\hline
\end{tabular}
\label{tab:result1}
\end{table*}

\begin{table*}
\centering
\caption{Summary of radiative transfer runs with different SiO fractional abundance
  profiles for dynamical simulation with molecular cooling and $\eta=3$.}
\begin{tabular}{l | l | l | l | l}
\hline
Profile & v$_{\rm j}$ [km s$^{-1}$] & $\delta^{-1}$ = v$_{\rm j}$/v$_{\rm s}$ & Peak Intensity [K] & $\Delta$v [km s$^{-1}$]\\ 
\hline\hline
Top Hat & 100.0 & 1.0 & & \\
Functional & 100.0 & 1.0 & & \\
Functional & 100.0 & $(1.0 + \eta^{-0.5})$ &
& \\
Functional & 100.0 & $(1.0 + \eta(z)^{-0.5})$ &
& \\
Functional & 125.0 & $(1.0 + \eta(z)^{-0.5})$ &
& \\
Functional & 150.0 & $(1.0 + \eta(z)^{-0.5})$ &
& \\
\hline
\end{tabular}
\label{tab:result2}
\end{table*}

\section{Results}
\subsection{Comparison of cooling prescription}
Importance of cooling prescriptions\\
Bulk of emission comes from cooling instabilities.
figure nos. 2 -- Similuated images, emission maps for eta = 10.
\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/jetplt1_1010.png}
 \caption{Jet Volume Density for different cooling modes with
   $\eta$ = 10 and $\beta$ = 10.}
\label{grid}
\end{figure*}

\begin{figure}
 \includegraphics[width=84mm]{\figpath/J2-1_mom0maps_cooling.png}
 \caption{A plot of the integrated SiO J2-1 emission from 5 models, 
each using a different method to calculate cooling and all with $\eta$=10 $\beta$=10.}
\label{cooling2-1} % This plots needs to be changed.
\end{figure}

% \subsection{Effects of changing $\eta$ and $\beta$}
% figure nos. 2 -- Simulated images describing lengths, collimation and
% emission for lower eta more stronger. 
% \begin{figure*}
%  \includegraphics[width=2.\columnwidth]{\figpath/jetplt2_diffetabeta.png}
%  \caption{Jet Volume Density for atomic cooling with $\eta$ = 10 but $\beta$ = 1 and
%    10 and that with tabulated cooling with $\beta$ = 10 but $\eta$ = 2
%    and 10.}
% \label{grid}
% \end{figure*}

\subsection{Molecular cooling and H$_2$ Chemistry}
In case of molecular outflows, the most appropiate form of cooling
prescribtion among the ones described in sect.~\ref{sec:chem} is
that which involves the evolution of H$_2$ chemistry along with
contributions to cooling from fixed fractions of other molecules like
CO, OH etc. First three panels of figure~\ref{fig:Hfracsim} shows the density of
various hydrogen species in the outflow for the simulation run with
$\eta$ = 3 and $\beta$ = 10. The jet is largely dominated by atmoic
and molecular hydrogen, however, the fraction of these species have
considerably changed from their initial values within the
jet. Ionzied hydrogen is mainly formed at the tip
of the bow as seen by the increase of fHII to 10\% from an initial value of
1\% within the jet.
%

The last two panels of figure~\ref{fig:Hfracsim}
show the temperature and mean molecular weight $\mu$. The
highest tempeature of $\sim$ 50000\,K is attained in our flow at the
tip of the bow shock. While the temperature on the edges (i.e.,interface between
jet and the ambient medium) is lower than 5000\,K. Also the relatively
weaker shocks formed due to knots do not heat up the material beyond
few 10$^{3}$\, K. The mean molecular weight, $\mu$, gives an indication
on which specie of hydrogen dominates in what regions of the flow. In
particular, a value of $\mu > 2.$ represents regions dominated by
molecular hydrogen, regions with lowest values of $\mu$ are where most
of the ionized hydrogen is present, while along the jet the value
$\mu$ is close to 1.3 suggesting an atomic jet. 
%

\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/molcjet_310.png}
 \caption{Fraction of hydrogen species in the run with molecular
   cooling having $\eta$ = 3 and $\beta$ = 10.}
\label{fig:Hfracsim}
\end{figure*}

The distribution of fractions of different hydrogen species along with
temperature suggests that there are essentially three regions where
chemistry is evolved due to shocks.
They are : (1) The tip of the jet , (2) The edges of the jet and
(3) intermediate knots. As the atomic jet propgates from the lower boundary 
into the cold molecular medium , it forms a strong shock resulting in
forming a density and temperature discontinuity. Such a jump in
dynamical quantities play a crucial role in evolution of
chemistry. For example, temperatures beyond few 1000\,K produced in the shocks could
disassociate the molecules and can also lead to ionization if
temperature reaches above 10$^{4}$\,K.

\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/molcjet_310_molform.png}
 \caption{Dependence of hydrogen fractions on the temperature at two
   points in the flow, viz. the interface of the knot with molecular
   medium and at the bow shock}
\label{fig:Hmolform}
\end{figure*}

\subsection{SiO Abundance profile}
figure nos. 2 -- Comparison of three profiles for SiO Abundance,
Contour maps for two transitions for top hat profile.
Effects of varing abundance profile ... Crucial ingredient in
modeling molecular bullets \\
\begin{figure}
 \includegraphics[width=1.\columnwidth]{\figpath/diffabunEM.png}
 \caption{Variation of 2-$>$1 SiO emission for runs with molecular
   cooling having $\eta$ = 3 and $\beta$ = 10 and different abundance profiles.}
\label{grid}
\end{figure}

\begin{figure}
 \includegraphics[width=1.\columnwidth]{\figpath/PV_diffprof.png}
 \caption{Contour maps of position-velocity diagram for the internal
   knots for 2-$>$1 SiO emission for runs with molecular
   cooling having $\eta$ = 3 and $\beta$ = 10 and different abundance
   profiles. The contours mark different levels of emission in Kelvins, viz.,
   0.2,0.6,1.0,1.4,1.8,2.0,3.0,4.0.}
\label{grid}
\end{figure}

\subsection{Spectra and PV diagrams}
figure nos. 1 -- one transition 3 different angles, Comparison of
spectra and line ratios  
Spectra and PV Digrams\\
Line - withs $\sim$ 15-25 km\,s$^{-1}$. Wedge shaped PV diagrams
a signature feature of jet driven molecular outflows. Counterparts in
observations.

\begin{figure}
 \includegraphics[width=1.\columnwidth]{\figpath/imshkvh_21_90_empvspec.png}
 \caption{Emission map ({\it top left}), PV diagram ({\it top right}
   and spectra ({\it bottom} for the
   2-$>$1 SiO transition for the run with molecular cooling with
   $\eta$ = 3 and fractional abunance of functional form is used
 along with $\delta < $ 1. The jet is assumed to be in the plane of
 sky implying an angle of inclination of $90^{\circ}$. The horizontal
 {\it white dashed} line marks the cuts where the spectra is taken
 while the vertical {\it magenta dashed} line represents the cut for
 the PV digram.} 
\label{empvspec90}
\end{figure}

\begin{figure}
 \includegraphics[width=1.\columnwidth]{\figpath/imshkvh_21_60_empvspec.png}
 \caption{Same as figure~\ref{empvspec90} but with angle of
   inclination of $60^{\circ}$.} 
\label{empvspec60}
\end{figure}

\begin{figure}
 \includegraphics[width=1.\columnwidth]{\figpath/imshkvh_21_45_empvspec.png}
 \caption{Same as figure~\ref{empvspec90} but with angle of
   inclination of $45^{\circ}$.} 
\label{empvspec45}
\end{figure}

\section{Discussion}
Here we compare the various results with SiO observation of young low
mass and high mass outflows.

\subsection{Line transitions and ratio}
Different line transition trace different components of SiO
emission. The low level transitions trace the region where the jet
interacts with molecular medium. The clumpy nature in its emission is
from the turbulent nature of its interaction. The higher transitions
trace a much more collimated region along with the terminal shock. 
As seen by \cite{Chandler:2001p14376} and \cite{Hirano:2006p14411}.
% \begin{figure*}
%  \includegraphics[width=2.\columnwidth]{\figpath/linejets_thnolog.png}
%  \caption{SiO emission from different line transitions for the run with molecular
%    cooling having $\eta$ = 3 and $\beta$ = 10 using top-hat fractional
%    abundance profile. }
% \label{grid}
% \end{figure*}

\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/imshkvhcontomaps.png}
 \caption{Contours of SiO emission for different line transitions
   obtained for the run with molecular cooling with $\eta$ = 3 using
   the functional form of the SiO fractional abundance with $\delta <$
   1. The contour colors represent different intensities in Kelvins, i.e,
   30.0({\it red}), 10.0({\it green}), 5.0({\it blue}), 1.0({\it
     magenta}), 0.5({\it cyan}), 0.1({\it black}).} 
\label{grid}
\end{figure*}

Advantage of non-LTE modeling over LVG \cite{Cabrit:2007p13804}, \cite{Lee:2008p13697}. 
Line ratios $\sim$ 1 can be explained well for low mass stars and 
\cite{Leurini:2013p13165} for high mass stars. \\
\begin{figure}
 \includegraphics[width=1.\columnwidth]{\figpath/LineRatio_imshkv60.png}
 \caption{{\em{Top}} Line profiles in SiO J = 2-1, 5-4 and 8-7 at one
   the inner knot for the molecular cooling run with $\eta$ = 3 and
   $\beta$ = 10 using top hat abundance profile. The profiles are
   obatined when the angle of inclination is 60$^{\circ}$ with respect
 to line of sight. {\em{Bottom}} Line temperature ratios T(8-7)/T(5-4)
and T(2-1)/T(5-4), as a function of velocity}
\label{grid}
\end{figure}

\subsection{Shock Kinematics}
C-type vs J-type shocks. \cite{Gueth:1998p14058} shows that SiO
emission has counterparts to H$_2$ emission and prefers the J-shock
formation of SiO.

\subsection{ALMA view}
ALMA view of the reference run and stress of applying our synthetic techniques
to study the molecular outflows in more details.\\
Figure 6. for ALMA --
